Title Goes Here

### 2024-04-17 analysis/paper/paper.qmd
stash

>>>>

```{r}
# method focus by year

method_foci <-
  read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0",
             sheet = "method")

method_foci_clean <-
  map(
    str_split(method_foci$`From the data`, ";"),
    ~.x %>%
      str_squish() %>%
      stri_remove_empty() %>%
      str_to_lower)

jobdata_method <-
  jobdata %>%
  select(methods_focus_of_position) %>%
  mutate(methods_focus_of_position = str_to_lower(methods_focus_of_position))

jobdata_method <-
  # add one column for each topic in our categories
  cbind(jobdata_method,
        setNames( lapply(method_foci$Category, function(x) x=NA),
                  method_foci$Category) )

for(i in 1:length(method_foci$Category)){

  this_method <- method_foci$Category[i]

  # create the pattern to search for
  x <- paste0(method_foci_clean[[i]], collapse = "|")

  # do the search through all the job ads for that pattern
  y <- str_detect(jobdata_method$methods_focus_of_position,
                  x)

  # assign back to our data frame in the appropriate location column
  jobdata_method[, this_method] <- y

jobdata_method_year <-
  jobdata %>%
  bind_cols(jobdata_method) %>%
  select(year_ad_posted,
         method_foci$Category) %>%
  pivot_longer(-year_ad_posted) %>%
  drop_na()

# how many times each topic mentioned?
jobdata_method_year %>%
  group_by(name) %>%
  summarise(n = sum(value)) %>%
  arrange(desc(n)) %>%
  ggplot() +
  aes(reorder(name, n),
      n)+
  geom_col() +
  xlab("") +
  theme_minimal() +
  coord_flip()

# explore trends over time. put a point on the max year
jobdata_method_year_tally <-
  jobdata_method_year %>%
  # exclude those with <20 ads
  filter(!name %in% c("Digital Archaeology",
                      "Pleistocene archaeology",
                      "Mesoamerican Archaeology",
                      "Biological anthropology",
                      "Archaeological theory",
                      "Evolutionary anthropology",
                      "North American archaeology"
  )) %>%
  group_by(year_ad_posted,
           name) %>%
  summarise(n = sum(value)) %>%
  mutate(prop = n / sum(n))

jobdata_method_year_tally_max <-
  jobdata_method_year_tally %>%
  group_by(
    name ) %>%
  filter(prop == max(prop))

ggplot() +
  geom_smooth(data =  jobdata_method_year_tally,
            aes(year_ad_posted,
                prop,
                group = name,
                colour = name),
            size = 3,
           span = 0.7,
            se = FALSE) +
  xlab("Year") +
  ylab("Proportion of all ads") +
  scale_colour_brewer(palette = "Dark2") +
  guides(colour = guide_legend("Method\nfocus",
                               label.position = "bottom")) +
    theme_minimal( base_size = 14) +
  theme(legend.position="bottom")

ggsave(here("analysis",
            "figures", 
            "fig-topic-focus-by-year.png"),
       bg ="white",
       h = 10, # experiment with h and w to get the right size and proportion
       w = 12,
       units = "in",
       dpi = 900) # make the image nice and crisp))
```
